Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2023-01-17 paper-1/manuscript.tex
Editing

Next, we turned to the likely cause of this difference, the non-target responses (Figure \ref{fig5}b), mapped into space as the average number of spikes fired in each neuron, $\R{N}{i}$. As expected, the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds. Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.

@@

Finally, we examined the cause of the reduced response in non-target trials. Going from the \msc{} sequence to the \dev{} sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation. Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response. We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and comparatively strongly to stimulation in B, and therefore more to non-target stimuli in \dev{} than \msc{}, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.

@@

To show this, we sought to isolate this subset of early responders to B. In a given network and stimulus assignment, we found, for each neuron, the minimum response time (i.e., time to first spike) across B trials in all sequences, and ranked neurons by this response time. The ranking in the sample network is shown in Figure \ref{fig6}a. We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site. We did notice that the early response appeared to gravitate towards the stimulus site for A, but confirmed that this was the case in all sequences separately, indicating a structural rather than dynamic cause (data not shown). We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons. In Figure \ref{fig6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\T{}{}$ at the start of B trials, $\Delta \T{B}{i}$. Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$. As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the \dev{} sequence, while the response of the later portion of the network decreased. Correspondingly, the TA levels of the early portion also tended to increase, in clear contrast to the later portion.

Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B. We first confirmed that $\Delta \R{N}{i}$ and $\Delta T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120). In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, Figure \ref{fig6}c) and the median contrast in $\T{}{}$ at the start of B trials ($\Delta \T{B}{}$, Figure \ref{fig6}d) within these two portions. We found that the response in the early portion was consistently greater in \dev{} than \msc{} (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d). $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).

Finally, to show that early $\T{B}{}$ was causal in reducing the response to B, we correlated the median contrast in $\T{}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in Figure \ref{fig6}e. Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.

@@

As noted previously, STD alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA. To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD. On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (see Figure \ref{fig7}a). However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (T = 2.66e+03, p = 0.00554).

@@

To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section. Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}b the full model equivalent of Figure \ref{fig6}c and d. The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence. Likewise, the correlation across excitatory neurons between these two contrasts remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992). This indicates that local suppression to B mediated by TA remained a key driver of the reduced response to non-target trials.

Comparing the non-target response reductions in Figure \ref{fig7}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%. Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials. Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}d. Correspondingly, $\T{A}{exc}$ was reduced in \dev{}, as shown in Figure \ref{fig7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).

@@

Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes. The primary cause for the greater target response, therefore, appeared to be the smaller non-target response, which was where we noted the first clear shift towards increased deviance detection.

@@

To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated the number of spikes in \dev{} target trials with $x_j$ at the start of later trials (cf. Equation \ref{eqn-xsyn}) across excitatory neurons. We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial. Similar behavior was also seen in the \msc{} sequence, as shown in Figure \ref{sfig-STD-correlation-decay}.

To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} responses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD. Unlike $\T{}{}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse. Therefore, we devised a measure $\D{}{}$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $\T{}{}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.

A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity. In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $\R{S}{j} = \frac{1}{3} \sum_{seq \in (std, dev, msc)} \R[seq]{S}{j}$ of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive. Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$. Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $\R{S}{j} k x_j^S$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials. Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
\begin{equation}
    \D{S}{i} = \sum_{j \in pre} \R{S}{j} k (1-x_j^S) \label{eqn-D}
\end{equation}
Notice that $\D{}{}$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $\T{}{}$, albeit more approximate.

@@

As noted above, we hypothesized that, in the full model, STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly. To show this, we calculated $\D{}{}$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences. The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast $\Delta \D{A}{i}$, which was negative throughout. Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both sequences. In the \msc{} sequence, we can attribute the difference between A and B primarily to different response patterns due to network structure, not dynamics: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network. However, in the \dev{} sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited. As a result, the contrast $\Delta \D{B}{i}$ was slightly reduced in the periphery, but strikingly elevated near the stimulation site, indicating greatly increased synaptic depression in \dev{}.

To show that $\D{}{}$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated $\Delta \D{A}{i}$ with $\Delta \R{A}{i}$ across neurons of the sample network. Notice that this mirrors the correlation shown in Figure \ref{fig4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$. As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected. We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120), though the median correlation coefficient $\rho$ was only -0.282. We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).

@@

To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing Figure \ref{fig6}e. As we show in Figure \ref{fig8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000). This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.

@@

To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli (lower $\R[dev]{B}{}$) by local suppression, and an increase in the response to target stimuli (higher $\R[dev]{A}{}$) by relief from global suppression, relative to \msc{}. Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$. Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.

We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude. To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}e, bottom) with the corresponding $\Delta \D{A}{}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV. Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression. Thus, this synergy allows the individually weak STD to exert a significantly larger effect when paired with the stronger TA.

>>>>

Visualizing the likely cause of this difference, the non-target responses, in the same fashion (Figure \ref{fig5}b), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected. Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.

@@

Finally, we examined the cause of the reduced response in non-target trials. Going from the \msc{} sequence to the \dev{} sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation. Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response. We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B, and therefore more to non-target stimuli in \dev{} than \msc{}, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.

@@

To show this, we sought to isolate this subset of early responders to B. In a given network and stimulus assignment, we found, for each neuron, the minimum response time (i.e., time to first spike) across B trials in all sequences, and ranked neurons by this response time. The ranking in the sample network is shown in Figure \ref{fig6}a. We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site. We did notice that the early response appeared to gravitate towards the stimulus site for A, but confirmed that this was the case in all sequences separately, indicating a structural rather than dynamic cause (data not shown). We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons. In Figure \ref{fig6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\T{}{}$ at the start of B trials, $\Delta \T{B}{i}$. Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$. As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red, left plot) increasing their response in the \dev{} sequence, while the response of the later portion of the network (blue, right plot) decreased. Correspondingly, the TA levels of the early portion also tended to increase, in clear contrast to the later portion.

Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B. We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120). In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, Figure \ref{fig6}c) and the median contrast in $\T{}{}$ at the start of B trials ($\Delta \T{B}{}$, Figure \ref{fig6}d) within these two portions. We found that the response in the early portion was consistently greater in \dev{} than \msc{} (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d). $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).

Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in Figure \ref{fig6}e. Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.

@@

As noted previously, STD alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA. To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD. On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (Figure \ref{fig7}a). However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (more than \msc{}: T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (less than \msc{}: T = 2.66e+03, p = 0.00554).

@@

To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section. Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}b the full model equivalent of Figure \ref{fig6}c and d. The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence. Likewise, the correlation across excitatory neurons between these two contrasts remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992). This indicates that local suppression to B mediated by TA in the early responders remained a key driver of the reduced response to non-target trials.

Comparing the non-target response reductions in Figure \ref{fig7}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%. Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials. Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}d. Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in Figure \ref{fig7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).

@@

Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes. The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (Figure \ref{fig7}c), which was where we noted the first clear shift towards increased deviance detection.

@@

To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated the number of spikes in \dev{} target trials with $x_j$ at the start of later trials across excitatory neurons. We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial. Similar behavior was also seen in the \msc{} sequence, as shown in Figure \ref{sfig-STD-correlation-decay}.

To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} responses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD. Unlike $\T{}{j}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse. Therefore, we devised a measure $\D{}{}$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $\T{}{}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.

A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity. In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $\R{S}{j} = \frac{1}{3} (\R[dev]{S}{j} + \R[msc]{S}{j} + \R[std]{S}{j})$ of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive. Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$. Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $\R{S}{j} x_j^S k$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials. Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
\begin{equation}
    \D{S}{i} = \sum_{j \in pre} \R{S}{j} (1-x_j^S) k \label{eqn-D}
\end{equation}
Notice that $\D{}{}$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $\T{}{}$, albeit more approximate.

@@

As noted above, we hypothesized that STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly. To show this, we calculated $\D{}{}$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences. The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast $\Delta \D{A}{i}$, which was negative throughout. Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both sequences. In the \msc{} sequence, we can attribute the difference between A and B primarily to different response patterns due to network structure, not dynamics: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network. In the \dev{} sequence, however, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited. As a result, the contrast $\Delta \D{B}{i}$ was slightly reduced in the periphery, but strikingly elevated near the stimulation site, indicating greatly increased synaptic depression in \dev{}.

To show that $\D{}{}$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated $\Delta \D{A}{i}$ with $\Delta \R{A}{i}$ across neurons of the sample network. Notice that this mirrors the correlation shown in Figure \ref{fig4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$. As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected. We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120). We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).

@@

To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing Figure \ref{fig6}e. As we show in Figure \ref{fig8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120). This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.

@@

To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli (lower $\R[dev]{B}{}$) by local suppression, and an increase in the response to target stimuli (higher $\R[dev]{A}{}$) by relief from global suppression, relative to \msc{}. Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower thresholds $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$. Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.

We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude. To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV. Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression. Thus, this synergy allows the individually weak STD to exert a significantly larger effect when paired with the stronger TA.
